---
title: "descriptives02"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
# knitr opts
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
# packages
library(tidyverse)
library(sf)
library(mapview)
library(leaflet)
library(leafem)
library(leafpop)
library(leafsync)
library(plotly)
library(miceadds)
library(shiny)
library(shinyWidgets)

processed   <- "/Volumes/Al-Hakem-II/Datasets/bks/bks"
days    <- readRDS(file.path(processed, "data/plato/days.Rda"))
#bks     <- readRDS(file.path(processed, "data/plato/daily-rides.Rda"))
sum_station <- readRDS(file.path(processed, "data/plato/sum-station.Rda")) %>%
  select(-starts_with("lag")) %>% # remove lag vars, not needed here.
  st_as_sf(coords = c('lng', 'lat'), na.fail = FALSE) # set as sf object
  st_crs(sum_station) <- 4326
sum_station_yr <- readRDS(file.path(processed, "data/plato/sum-station-yr.Rda")) %>%
select(-starts_with("lag")) %>% # remove lag vars, not needed here.
  st_as_sf(coords = c('lng', 'lat'), na.fail = FALSE) # set as sf object
  st_crs(sum_station_yr) <- 4326

# document/package settings
theme_set(theme_minimal()) # set ggplot theme
mapviewOptions(fgb = T) # set to false for greater performance?
alpha = 0.2
```

Thanks to RColorBrewer for providing the color scales.

# Days

```{r}
inputPanel(
  dateRangeInput('date1', "Date",
                 startview = 'year', start = '2010-01-01', end = '2020-12-31'),
  selectInput('y1', "Y Axis", 
              choices = c("nrides",
                          "dur_med",
                          "dur_ineq"),
              selected = "nrides",  multiple = FALSE, selectize = TRUE)
) # note this must be "extracted" to be translated into bks data since not in date format



in.yr    <- reactive({lubridate::year(input$date1)})
#in.yrday <- reactive({lubridate::yday(input$date1)})

data <- reactive({
  filter(days, year >= in.yr()[1], year <= in.yr()[2])
})
renderPlot({
  ggplot(data(), aes(day_of_yr, input$y1, color=as.factor(year))) + 
   geom_point(alpha=0.6) + geom_smooth()
}) 




```

# Stations

## Yearly Departures 

```{r, echo=FALSE}
inputPanel(
  sliderInput('year1', "Year", 2010, 2020, 2010, sep = '')
)

data1 <- reactive({ filter(sum_station_yr, year == input$year1) })
m1 <- reactive({
  mapview(data1(), zcol = "departures",
          at = c(0,1000,2000,5000,10000,20000,100000)
          )
  })

renderLeaflet({
    m1()@map
})


```

## Inputs and Outputs

You can embed Shiny inputs and outputs in your document. Outputs are automatically updated whenever inputs change. This demonstrates how a standard R plot can be made interactive by wrapping it in the Shiny `renderPlot` function. The `selectInput` and `sliderInput` functions create the input widgets used to drive the plot.

```{r eruptions, echo=FALSE}
inputPanel(
  selectInput("n_breaks", label = "Number of bins:",
              choices = c(10, 20, 35, 50), selected = 20),
  
  sliderInput("bw_adjust", label = "Bandwidth adjustment:",
              min = 0.2, max = 2, value = 1, step = 0.2)
)

renderPlot({
  hist(faithful$eruptions, probability = TRUE, breaks = as.numeric(input$n_breaks),
       xlab = "Duration (minutes)", main = "Geyser eruption duration")
  
  dens <- density(faithful$eruptions, adjust = input$bw_adjust)
  lines(dens, col = "blue")
})
```

Note the use of the `height` parameter to determine how much vertical space the embedded application should occupy.
