---
title: "descriptives02"
output: 
  html_document:
    toc: true
    runtime: shiny
---

```{r setup, include=FALSE}
# knitr opts
# fyi: plotly::schema()
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
# packages
library(tidyverse)
library(sf)
library(mapview)
library(leaflet)
library(leafem)
library(leafpop)
library(leafsync)
library(plotly)
library(shiny)
library(shinyWidgets)
library(lubridate)
library(timetk)
library(RColorBrewer)

reactlog::reactlog_enable()

processed   <- "/Volumes/Al-Hakem-II/Datasets/bks/bks"
days    <- readRDS(file.path(processed, "data/plato/days.Rda")) %>%
  mutate(dur_med = round((dur_med/60),1)) %>% # change to rounded minutes
  rename(maxtemp = tempmax) # rename for variable order
#bks     <- readRDS(file.path(processed, "data/plato/daily-rides.Rda"))
sum_station <- readRDS(file.path(processed, "data/plato/sum-station.Rda")) %>%
  mutate(dur_med = round((dur_med/60),1)) %>%
  select(-starts_with("lag")) # remove lag vars, not needed here.
  
sum_station_yr <- readRDS(file.path(processed, "data/plato/sum-station-yr.Rda")) %>%
select(-starts_with("lag"))  # remove lag vars, not needed here.


# document/package settings
theme_set(theme_minimal()) # set ggplot theme
mapviewOptions(fgb = T) # set to false for greater performance?
alpha = 0.2
```

Thanks to RColorBrewer for providing the color scales.

# Days

The daily data is plotted in blue, along with optional weather data from NOAA. Date ranges can be subset by dragging either on the *Range Selector* below or on the plot region itself. Double-clicking resets.

```{r}

inputPanel(
  verticalLayout(
    tags$h4("Bikeshare Data"),
    pickerInput('y1', 
              choices = c("Total Daily Rides"      =  "nrides",
                          "Median Ride Duration" =  "dur_med",
                          "Duration Inequity"    =  "dur_ineq"),
              selected = "nrides",  multiple = FALSE, 
              options = pickerOptions(mobile = T))),
  verticalLayout(
      tags$h5("Options"), # spacing
      prettySwitch('y1.weather', "Show Temperature",
               value = FALSE, slim = T, fill = T, inline = T),
      prettySwitch('y1.precip', "Show Precipitation",
               value = FALSE, slim = T, fill = T, inline = T )),
  verticalLayout(
    tags$br(),tags$br(),
    prettySwitch('y1.tempfill', "Use Temperature as fill color",
                 value = FALSE, slim = T, fill = T, inline = T),
    awesomeCheckbox('y1.fahr', "Use ℉", value = FALSE, status = "primary"),
    actionButton('go', "Update")
  )
) # end vert layout and input panel


# data wrangling
# rolling average, 30 days
roll_av_30 <- timetk::slidify(.f = ~ mean(., na.rm=T), .period = 30, .align = 'center', .partial = TRUE)
roll_av_3 <- timetk::slidify(.f = ~ mean(., na.rm=T), .period = 3, .align = 'center', .partial = TRUE)

# midpoints
mid <- reactive({
  if (input$y1.fahr) {
    60 # far midpoint 
  } else {
    15 # celsius
  }
})

min <- reactive({
  if (input$y1.fahr) {
    0  # far min 
  } else {
    -12 # celsius
  }
})

max <- reactive({
  if (input$y1.fahr) {
    110  # far min 
  } else {
    42 # celsius
  }
})

ticks <- reactive({
  if (input$y1.fahr) {
    c(20,40,60,80,100)  # far min 
  } else {
    c(0,10,20,30,40) # celsius
  }
})

#name
name <- reactive({
  case_when(
        input$y1 == "nrides"     ~ "Daily Rides",
        input$y1 == "dur_med"    ~ "Median Duration (min)",
        input$y1 == "dur_ineq"   ~ "Duration Inequity Index",
      )
})

d <- reactive({
  if (input$y1.fahr) {
    days %>% dplyr::ungroup() %>%
      mutate(maxtemp = round(((9/5)*maxtemp)+32),2)
  } else {
    days %>% dplyr::ungroup()
  }
})

data1main <- reactive({
  d() %>% 
  select(date, input$y1, maxtemp) %>%
  mutate(
    rolling_av_30 = case_when( # round only duration_ineq to the hundreths
      input$y1 == "dur_ineq" ~  round(roll_av_30(.data[[input$y1]]), 2),
      input$y1 != "dur_ineq" ~  round(roll_av_30(.data[[input$y1]]), 1)))
})


data1w <- reactive({
  d() %>% 
  select(date, maxtemp) %>%
  mutate(rolling_av_30_temp =  round(roll_av_30(.data[["maxtemp"]])))
})
  
data1p <- reactive({
  d() %>% 
  select(date, precip)
})

# graph
ay <- reactive({
  list(
    zeroline = F, showgrid = F,
    tickfont=list(color=brewer.pal(5,'Purples')[5]), overlaying='y',
       side='right', 
       title=list(text=paste('Observed Daily High',
                             if (input$y1.fahr){'℉'} else{'℃'}),  
                  font=list(color=brewer.pal(5,'Purples')[5])),
       visible=input$y1.weather)
}) 
by <- reactive({
  list(zeroline = F, showgrid = F,
       tickfont=list(color=brewer.pal(5,'Greys')[4]), overlaying='y', side='right', title=list(
    text='Observed Precipitation (mm)', standoff=35), visible=input$y1.precip)
})
  
# make the graph, depending on the go, button

  p1 <- eventReactive(input$go, {
    withProgress(message = "Loading", # start here.
                 
    plot_ly() %>%
    # temp data
    add_trace(data = data1w(), type='scatter', mode = 'markers', name = "Daily Temp.",
              x = ~date, y = ~maxtemp, 
              marker = list(color = brewer.pal(5,'Purples')[3],
                            opacity = 0.8,
                            size = 5), 
              opacity = 0.8,
              visible = input$y1.weather,
              legendgroup = "temp",
              hovertemplate = paste0("<b>%{y:,1f}°</b><br>",
                                   "%{x}",
                                   "<extra></extra>"),
              yaxis = "y2", alpha = 1 ) %>% #directs to layout.yaxis2
    # rolling av temp
    add_trace(data = data1w(), type='scatter', mode = 'lines', name = "30-day Av",
              x = ~date, y = ~rolling_av_30_temp, 
              line = list(color = brewer.pal(5,'Purples')[5]), 
              opacity = 0.9,
              visible = input$y1.weather,
              legendgroup = "temp",
              hovertemplate = paste0("<b>%{y:,1f}°</b><br>",
                                   "%{x}",
                                   "<extra></extra>"),
              yaxis = "y2", alpha = 1 ) %>%
    # precip
    add_trace(data = data1p(), type = 'bar', x = ~date, y = ~precip,
              marker = list(color = brewer.pal(5,'Greys')[3], opacity = 0.6),
              name="Precipitation",
              legendgroup = "precip",
              marker=list(line=list(width=1.5)), visible = input$y1.precip,
              hovertemplate = paste0("<b>%{y:1f} mm</b><br>",
                                   "%{x}",
                                   "<extra></extra>"),
              yaxis = "y3", alpha = 1 ) %>%
    # main data 
    add_trace(data = data1main(),
              type = 'scatter', mode = 'markers', name = name(),
              x = ~date, y = ~eval(as.symbol(input$y1)),
              marker = list(
                color = if (input$y1.tempfill) {~maxtemp} else {brewer.pal(5,'Blues')[3]}, 
                opacity = 0.9,
                size = 5,
                showscale = input$y1.tempfill,
                cmin = min(),
                cmid = mid(),
                cmax = max(),
                colorscale = "RdBu",
                colorbar = list(
                    title = list(text="Observed <br>Daily High", font=list(size=14)),
                    x = 1.17, y=0.6, len = 1, thickness=25,
                    tickvals=ticks(), tickmode = 'array',
                    ticksuffix=paste(if (input$y1.fahr){'℉'} else{'℃'})
                  )),
            legendgroup = "bks",
            hovertemplate = if_else( input$y1.tempfill,
              false = paste0("<b>%{y:,2f}</b><br>",
                            "%{x}",
                            "<extra></extra>"),
              true = paste0("<b>%{y:,2f}</b><br>",
                            "%{marker.color:1f}",if (input$y1.fahr){'℉'} else{'℃'},
                            "<br>%{x}",
                            "<extra></extra>"))
            ) %>%
    # rolling av: main data
    add_trace(data = data1main(),
              type = 'scatter', mode = 'lines', name = paste0("30-day Av"),
              x = ~date, y = ~rolling_av_30, 
              line = list(color = brewer.pal(5,'Blues')[4]), 
            legendgroup = "bks",
            hovertemplate = paste0("<b>%{y:,1f}</b><br>",
                                   "%{x}",
                                   "<extra></extra>")) %>%
  layout( 
    title = list(text=""),
    yaxis2= ay(),
    yaxis3= by(),
    yaxis = list(
      showgrid = F,
      title = list(text=paste("<b>",name(),"</b>"), font=list(size=18))), 
    xaxis = list(
      title = list(text="<b>Date Range Selector</b>", font=list(size=16)),
      rangeslider = list(autorange=TRUE, thickness = 0.15),
      rangeselector = list(
        x = 0.02, y = 1.1,
        buttons = list(
        list(count=3, label="3 yr", step='year', stepmode='backward'),
        list(count=1, label="1 yr", step='year', stepmode='backward'),
        list(count=3, label='3 mo', step='month',stepmode='backward'),
        list(step = 'all')))),
    font = list(family="arial"),
    legend = list(orientation='h', y=1.30, x=0.95, xanchor = "auto"),
    margin = list(l=80, r=100)
  ) %>% # end layout
    config(displayModeBar = FALSE)
    
  ) # end withprogress

}, ignoreNULL=FALSE, ignoreInit = FALSE)
  



# render the graph
renderPlotly({p1()})

 

```





### Terms and notes

**Number of Rides**: The total number of rides across the system that originated on a specific day.

**Median Ride Duration**: A measure of how long the 'average' ride lasted, converted to minutes. Mathematically, the median is used in lieu of average to avoid upward-skewing from extremely long rides (such as those that forgot to properly dock the bike).

**Duration Inequity**: An index from 0-1 that measures the "even-ness" of the daily ride duration distribution. Figures closer to 1 indicate that rides for that day were concentrated among just a few ride-length times, whereas figures closer to 0 indicate that rides, on aggregate, were evenly distributed across all ride-length times. This is simply a transport-oriented application of the GINI Index.

**Precipitation**: (1-2 sentence methodology explanationx)

# Stations

## Yearly Departures

```{r, echo=FALSE}
inputPanel(
  sliderInput('year2', "Year", 2010, 2020, 2010, sep = '', ticks = FALSE),
  selectInput('y2', "Y Axis", 
              choices = c("Number of Rides"      =  "departures",
                          "Median Ride Duration" =  "dur_med",
                          "Departure Inequity"   =  "dep_ineq",
                          "Near Metro"           =  "metro",
                          "Percent of Rides to Metro" = "metro_end_pct",
                          "Member Percent"       = "member_pct",
                          "Net Flow"             = "net_flow",
                          "Net Median Duration"  = "net_median_dur"),
              selected = "departures",  multiple = FALSE, selectize = TRUE),
  materialSwitch('scaled', "Scaled", value = FALSE)
)

pal2  <- reactive({
  if (input$scaled == FALSE) {
    colorQuantile("YlOrRd", sum_station_yr$input$y2, 5)
  } else {
    colorNumeric("YlOrRd", sum_station_yr$input$y2)
  }
  })
data2 <- reactive({ filter(sum_station_yr, year == input$year2) })

m2 <- reactive({
  leaflet(data = data2(), height = 800) %>%
  addTiles() %>%
  addCircleMarkers(color = ~pal2()(eval(as.symbol(input$y2))),
                   fillColor = ~pal2()(eval(as.symbol(input$y2))),
                   radius=7, stroke = TRUE, weight = 2,
                   opacity = 0.9, fillOpacity = 0.4, 
                   label = ~eval(as.symbol(input$y2)),
                   popup = ~name_bks_st) # add legend later, lol
})
  
renderLeaflet({m2()})



# data1 <- reactive({ filter(sum_station_yr, year == input$year1) })
# m1 <- reactive({
#   mapview(data1(), zcol = "departures",
#           at = c(0,1000,2000,5000,10000,20000,100000)
#           )
#   })
# 
# renderLeaflet({
#     m1()@map
# })


```

## Daily Departures

Note the interpretation is in terms of average or summed station-days over the time period specified.

```{r, echo=FALSE}
inputPanel(
  dateRangeInput('dt3a',"Date A","2010-01-01","2012-12-31", startview = "year"),
  dateRangeInput('dt3b',"Date B","2017-01-01","2020-12-31", startview = "year"),
  selectInput('y3', "Y Axis", 
              choices = c("Number of Rides"      =  "departures",
                          "Median Ride Duration" =  "dur_med",
                          "Departure Inequity"   =  "dep_ineq",
                          "Percent of Rides to Metro" = "metro_end_pct",
                          "Member Percent"       = "member_pct",
                          "Net Flow"             = "net_flow"),
              selected = "departures",  multiple = FALSE, selectize = TRUE),
  #materialSwitch('scaled3', "Scaled", value = FALSE),
  sliderInput('n3', 'N Colors', 2, 10, 5, width = 70),
  selectInput('name3', 'Colors', choices = hcl.pals("sequential")),
  actionButton('go', "Go")
)

# summarize/filter dataframe based on date range.
int3a  <- reactive({ interval(start = input$dt3a[1], end = input$dt3a[2]) }) 
data3a <- reactive({ 
  sum_station %>% filter(date %within% int3a() ) %>%
    dplyr::ungroup() %>% group_by(id_station, name_bks, lng, lat) %>%
    summarise(
      departures    = round(mean(departures, na.rm=T), 1),
      dur_med       = round(median(dur_med, na.rm=T), 1),
      dep_ineq      = round(mean(dep_ineq, na.rm=T), 3),
      metro_end_pct = round(mean(metro_end_pct, na.rm=T), 3),
      member_pct    = round(mean(member_pct, na.rm=T), 3),
      net_flow      = sum(net_flow)
    )
  })

int3b  <- reactive({ interval(start = input$dt3b[1], end = input$dt3b[2]) }) 
data3b <- reactive({ 
  sum_station %>% filter(date %within% int3b() ) %>%
    dplyr::ungroup() %>% group_by(id_station, name_bks, lng, lat) %>%
    summarise(
      departures    = round(mean(departures, na.rm=T), 1),
      dur_med       = round(median(dur_med, na.rm=T), 1),
      dep_ineq      = round(mean(dep_ineq, na.rm=T), 3),
      metro_end_pct = round(mean(metro_end_pct, na.rm=T), 3),
      member_pct    = round(mean(member_pct, na.rm=T), 3),
      net_flow      = sum(net_flow)
    )
  })


# pal3  <- reactive({ 
#   if (input$scaled3 == FALSE) {
#     colorQuantile("YlOrRd", sum_station$input$y3, 5)
#   } else {
#     colorNumeric("YlOrRd", sum_station$input$y3)
#   }
#   })
# 
# m3a <- reactive({
#   leaflet(data = data3a(), height = 800) %>%
#   addTiles() %>%
#   addCircleMarkers(color = ~pal3()(eval(as.symbol(input$y3))),
#                    fillColor = ~pal3()(eval(as.symbol(input$y3))),
#                    radius=7, stroke = TRUE, weight = 2,
#                    opacity = 0.9, fillOpacity = 0.4, 
#                    label = ~eval(as.symbol(input$y3)),
#                    popup = ~name_bks) # add legend later, lol
# })
# 
# m3b <- reactive({
#   leaflet(data = data3b(), height = 800) %>%
#   addTiles() %>%
#   addCircleMarkers(color = ~pal3()(eval(as.symbol(input$y3))),
#                    fillColor = ~pal3()(eval(as.symbol(input$y3))),
#                    radius=7, stroke = TRUE, weight = 2,
#                    opacity = 0.9, fillOpacity = 0.4, 
#                    label = ~eval(as.symbol(input$y3)),
#                    popup = ~name_bks) # add legend later, lol
# })

n  <- reactive({input$n3})
nm <- reactive({input$name3})


mv1 <- reactive({
  data3a() %>%  st_as_sf(coords=c("lng","lat"), na.fail=F) %>%
  mapview(., zcol = as.character(input$y3))
})
mv2 <- reactive({
  data3b() %>%  st_as_sf(coords=c("lng","lat"), na.fail=F) %>%
  mapview(., zcol = as.character(input$y3))
})
syncmap <- eventReactive(input$go, { 
  
  mapviewOptions( # change/set global options
  vector.palette = hcl.colors(input$n3, palette = input$name3)
  )
  
  sync( mv1(), mv2() )
  })



# renderLeaflet({ m3a() })
# renderLeaflet({ m3b() })
renderUI({syncmap()})
# 
# renderLeaflet({mv1()@map})
# renderLeaflet({mv2()@map})



```
